Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-11 paper.tex
increase image size (basic protocol) as it does no harm positioning-wise
change my email address
cleanup: removed old (commented out) text snippets - only in "my" parts

\newcommand{\symb}[1]{\makebox{\it #1}} 
\newcommand\ie{i.e.\!\,, }
\newcommand\rel{Re${\cal{L}}$}
%\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
%\newcommand*{\TODOfn}[2][noteC]{\TODO[#1]{[\footnote{\TODO[#1]{#2}}]}}
%\newcommand*{\TODOref}[2][todoC]{\TODOfn[#1]{\textbf{BIB-REF}: #2}}
\newcommand\ttand{\mbox{{$\land$}}}
\newcommand\ttor{\mbox{{$\lor$}}}
\newcommand\ttcap{\mbox{{$\cap$}}}
\newcommand\ttcup{\mbox{{$\cup$}}}
\newcommand\ttfun{\mbox{{$\Rightarrow$}}}
\newcommand\ttmimp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttimp{\mbox{{$\longrightarrow$}}}
\newcommand\ttequiv{\mbox{{$\equiv$}}}
\newcommand\ttexists{\mbox{{$\exists$}}}
\newcommand\ttforall{\mbox{{$\forall$}}}
\newcommand\ttneg{\mbox{{$\neg$}}}
\newcommand\ttneq{\mbox{{$\neq$}}}
\newcommand\ttin{\mbox{{$\in$}}}
\newcommand\ttnin{\mbox{{$\notin$}}}
\newcommand\ttImp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttlam{\mbox{\( \lambda \)}}
\newcommand\tttimes{\mbox{\( \times \)}}
\newcommand\ttlbrack{\mbox{\(\llbracket\)}}
\newcommand\ttrbrack{\mbox{\( \rrbracket \)}}
\newcommand\noie{\textit{i.e.},\xspace}
\newcommand\eg{~\textit{e.g.},\xspace}
\newcommand\noeg{\textit{e.g.},\xspace}
\newcommand\ttatI{\mbox{\( @_G \)}}
\newcommand\ttto[1]{\mbox{{$\to^{#1}$}}}
\newcommand\ttleq{\mbox{{$\le$}}}
\newcommand\ttrelI{\mbox{{$\to_{i}$}}}
\newcommand\ttrelIstar{\mbox{{$\to^*$}}}
\newcommand\ttrel[1]{\mbox{{$\to_{#1}$}}}
\newcommand\ttrelstar[1]{\mbox{{$\to_{#1}^*$}}}
\newcommand\ttalpha{\mbox{{$\alpha$}}}
\newcommand\ttsupseteq{\mbox{{$\supseteq$}}}
\newcommand\ttupdownarrow{\mbox{{$\Updownarrow$}}}
\newcommand\ttbigcap{\mbox{{$\bigcap$}}}
\newcommand\tttau{\mbox{{$\tau$}}}
\newcommand\ttsubseteq{\mbox{{$\subseteq$}}}
\newcommand\ttf{\mbox{{$f$}}}
\newcommand\ttvdash{\mbox{{$\vdash$}}}
\newcommand\ttref[1]{\mbox{\(\sqsubseteq_{#1}\)}}
\newcommand\ttrefV{\mbox{{$\sqsubseteq_V$}}}
\newcommand{\ttcalN}[1]{\mbox{{${\mathcal{N}}_{\texttt{#1}}$}}} 
\newcommand\ttNatt{\mbox{{$\mathcal{N}$}}}
\newcommand\ttattand[1]{\mbox{{$\oplus_{\wedge}^{#1}$}}}
\newcommand\ttattor[1]{\mbox{{$\oplus_{\vee}^{#1}$}}}
\newcommand\ttsigma{\mbox{{$\sigma$}}}
\newcommand\ttmref[1]{\mbox{{$\sqsubseteq_{#1}$}}}
\newcommand\ttmeref{\ttmref{\mathcal{E}}}
\newcommand\ttsn{\mbox{{\texttt{s}$_0$}}}
\newcommand\ttsigmap{\mbox{{$\sigma'$}}}
\newcommand\ttecal{\mbox{$\mathcal{E}$}}
%\newcommand\ttimg{\mbox{$\triangleleft$}}
\newcommand\ttimg{\mbox{\texttt{`}}}

@@

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bialut@gmail.com}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the recently published German
Corona-virus warning app. The app supports breaking infection chains by informing users
whether they have been in close contact to an infected person. The app has a decentralized
architecture that supports anonymity of users.
We provide a formal model of the existing app with the Isabelle Infrastructure framework
to show up some natural attacks in a very abstract model. We then use the security
refinement process of the Isabelle Infrastructure framework to highlight how the use of
continuously changing ephemeral ids improves the anonymity.
\end{abstract}

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%The Corona-virus app has been produced based on a sophisticated security
%concept conceived by experts in the field %to our knowledge, no formal verification
%%
%and has strong claims with regard to mathematical support:
%
%``Such a design builds on strong, mathematically provable support for privacy and data protection goals [...]'' \TODOref{DP-3T Whitepaper p. 2} \smallskip\\
%  \NOTE{DP-3T on their design and how wisely chosen it is ;) We have to use this! Something along the l%ines of:}
%
However, there has, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or to what extent trade offs have to be made due to inherent vulnerabilities.
The protocol implemented by the framework CWA resembles only the most basic
of the three protocols proposed by the DP-3T project.
Hence, a formal verified model that stresses the impact or limits of certain security measures
might give more weight to appeals like \cite{dp3t:impl}
%to Apple and Google
to adopt the more sophisticated protocols.
%
%Thereby, we believe
%that our current work is useful to increase the trust in the Corona-virus app necessary for
%its wide adoption which in turn is crucial for it to be efficient.

@@

There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
The debate among %emerging between
advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
%As first step, both attacks start with deanonymizing Ephemeral Ids,
%i.\,e. linking them the particular devices.
%Both attacks rely on deanonymisation of Ephemeral Ids,
%i.\,e. linking ids to particular devices.

\subsubsection{Basic DP-3T architecture}
%All DP-3T protocols roughly work the same:
%\subsubsection{Basic DP-3T protocol: How does it work?}
%  How does it relate to ENF/CWA? (broad strokes)}
%
Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

@@

In the most simple (and insecure) protocol proposed by DP-3T %\cite{dp3t:wp}[p14ff]
this basically translates into publishing the daily seeds used to derive EphIDs from.
The protocol implemented by ENF and,
hence, CWA adopts this low-cost design \cite{cwa:arch}. DP-3T proposes two other, more sophisticated protocols that improve
privacy and data protection properties to different degrees but are more costly to implement.
%Aside from additional security measures,
%\TODOfn[styC]{In short: I want to name it but don't talk about it (since we're not really concerned with% this level of detail). But this sounds just sh...}
%like signing content the server provides as a general rule, the protocol implemented by ENF and,
%hence, CWA follows this low-cost design.\TODOref{CWA solution architecture ($\rightarrow$ github:CWA); probably ENF specification} DP-3T proposes two other, more sophisticated protocols that improve
%privacy and data protection properties to different degrees but are more costly.
%%TODOfn[styC]{``complicated'' sounds lame but ``costly'' is, perhaps, to simple/literal} to set up.
%\pagebreak

Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
\begin{figure}[htb]
\vspace{-.5cm}
  \begin{center}
    \includegraphics[width=4in]{DP-3T_basic_protocol}
  \end{center}
\vspace{-.5cm}
  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
  \label{fig:dp3tprot}
\vspace{-.5cm}
\end{figure}

%\subsubsection{Attack scenarios}
%There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
%The debate among %emerging between
%advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

%In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).
%%As first step, both attacks start with deanonymizing Ephemeral Ids,
%%i.\,e. linking them the particular devices.
%Both attacks rely on deanonymisation of Ephemeral Ids,
%i.\,e. linking ids to particular devices.

>>>>

\newcommand{\symb}[1]{\makebox{\it #1}} 
\newcommand\ie{i.e.\!\,, }
\newcommand\rel{Re${\cal{L}}$}
%\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand\ttand{\mbox{{$\land$}}}
\newcommand\ttor{\mbox{{$\lor$}}}
\newcommand\ttcap{\mbox{{$\cap$}}}
\newcommand\ttcup{\mbox{{$\cup$}}}
\newcommand\ttfun{\mbox{{$\Rightarrow$}}}
\newcommand\ttmimp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttimp{\mbox{{$\longrightarrow$}}}
\newcommand\ttequiv{\mbox{{$\equiv$}}}
\newcommand\ttexists{\mbox{{$\exists$}}}
\newcommand\ttforall{\mbox{{$\forall$}}}
\newcommand\ttneg{\mbox{{$\neg$}}}
\newcommand\ttneq{\mbox{{$\neq$}}}
\newcommand\ttin{\mbox{{$\in$}}}
\newcommand\ttnin{\mbox{{$\notin$}}}
\newcommand\ttImp{\mbox{{$\Longrightarrow$}}}
\newcommand\ttlam{\mbox{\( \lambda \)}}
\newcommand\tttimes{\mbox{\( \times \)}}
\newcommand\ttlbrack{\mbox{\(\llbracket\)}}
\newcommand\ttrbrack{\mbox{\( \rrbracket \)}}
\newcommand\noie{\textit{i.e.},\xspace}
\newcommand\eg{~\textit{e.g.},\xspace}
\newcommand\noeg{\textit{e.g.},\xspace}
\newcommand\ttatI{\mbox{\( @_G \)}}
\newcommand\ttto[1]{\mbox{{$\to^{#1}$}}}
\newcommand\ttleq{\mbox{{$\le$}}}
\newcommand\ttrelI{\mbox{{$\to_{i}$}}}
\newcommand\ttrelIstar{\mbox{{$\to^*$}}}
\newcommand\ttrel[1]{\mbox{{$\to_{#1}$}}}
\newcommand\ttrelstar[1]{\mbox{{$\to_{#1}^*$}}}
\newcommand\ttalpha{\mbox{{$\alpha$}}}
\newcommand\ttsupseteq{\mbox{{$\supseteq$}}}
\newcommand\ttupdownarrow{\mbox{{$\Updownarrow$}}}
\newcommand\ttbigcap{\mbox{{$\bigcap$}}}
\newcommand\tttau{\mbox{{$\tau$}}}
\newcommand\ttsubseteq{\mbox{{$\subseteq$}}}
\newcommand\ttf{\mbox{{$f$}}}
\newcommand\ttvdash{\mbox{{$\vdash$}}}
\newcommand\ttref[1]{\mbox{\(\sqsubseteq_{#1}\)}}
\newcommand\ttrefV{\mbox{{$\sqsubseteq_V$}}}
\newcommand{\ttcalN}[1]{\mbox{{${\mathcal{N}}_{\texttt{#1}}$}}} 
\newcommand\ttNatt{\mbox{{$\mathcal{N}$}}}
\newcommand\ttattand[1]{\mbox{{$\oplus_{\wedge}^{#1}$}}}
\newcommand\ttattor[1]{\mbox{{$\oplus_{\vee}^{#1}$}}}
\newcommand\ttsigma{\mbox{{$\sigma$}}}
\newcommand\ttmref[1]{\mbox{{$\sqsubseteq_{#1}$}}}
\newcommand\ttmeref{\ttmref{\mathcal{E}}}
\newcommand\ttsn{\mbox{{\texttt{s}$_0$}}}
\newcommand\ttsigmap{\mbox{{$\sigma'$}}}
\newcommand\ttecal{\mbox{$\mathcal{E}$}}
%\newcommand\ttimg{\mbox{$\triangleleft$}}
\newcommand\ttimg{\mbox{\texttt{`}}}

@@

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bianca.lutz@campus.tu-berlin.de}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the recently published German
Corona-virus warning app. The app supports breaking infection chains by informing users
whether they have been in close contact to an infected person. The app has a decentralized
architecture that supports anonymity of users.
We provide a formal model of the existing app with the Isabelle Infrastructure framework
to show up some natural attacks in a very abstract model. We then use the security
refinement process of the Isabelle Infrastructure framework to highlight how the use of
continuously changing ephemeral ids improves the anonymity.
\end{abstract}

@@

The app now in use in Germany has been developed based on a protocol proposed by the DP-3T project \cite{dp3t:github}:
A sophisticated security concept
conceived by experts in the field that has strong claims with regard to mathematical support \cite{dp3t:wp} [p2].
%
However, there has, as of yet, to our knowledge no formal verification been involved.
Even if a ``post-production'' formal specification seems pointless, 
it allows to reveal weak points of the architecture, show that the measures that have
been conceived are suitable to cover those weak points,
or to what extent trade offs have to be made due to inherent vulnerabilities.
The protocol implemented by the framework CWA resembles only the most basic
of the three protocols proposed by the DP-3T project.
Hence, a formal verified model that stresses the impact or limits of certain security measures
might give more weight to appeals like \cite{dp3t:impl}
to adopt the more sophisticated protocols.

@@

There is a variety of noteworthy privacy and security issues. %that deserve formal consideration (and verification).
The debate among advocates of centralized architectures and those in favor of a decentralized approach in particular yields a lot of interesting material detailing different attacks and possible mitigation strategies: \cite{pepppt:dp3tana}, \cite{pepppt:dp3tresp}, \cite{dp3t:psre}.

In terms of attack scenarios, we focus on, what might be classified as deanonymisation attacks: Tracking a device (see \cite{dp3t:psre}[p9], \cite{pepppt:dp3tana}[p8]) and identifying infected individuals (\cite{dp3t:psre}[p5], \cite{pepppt:dp3tana}[p9]).

\subsubsection{Basic DP-3T architecture}
Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

@@

In the most simple (and insecure) protocol proposed by DP-3T %\cite{dp3t:wp}[p14ff]
this basically translates into publishing the daily seeds used to derive EphIDs from.
The protocol implemented by ENF and,
hence, CWA adopts this low-cost design \cite{cwa:arch}. DP-3T proposes two other, more sophisticated protocols that improve
privacy and data protection properties to different degrees but are more costly to implement.
%
Figure \ref{fig:dp3tprot} illustrates the basic system architecture
along with some of the mitigation measures either implemented in CWA or proposed by DP-3T.
\begin{figure}[htb]
\vspace{-.5cm}
  \begin{center}
    \includegraphics[width=5in]{DP-3T_basic_protocol}
  \end{center}
\vspace{-.5cm}
  \caption{Decentralized privacy-preserving proximity tracing protocol of CWA}
  \label{fig:dp3tprot}
\vspace{-.5cm}
\end{figure}
